\documentclass[pal,wide,12pt]{pajarticle}

\usepackage{pajmath}

\newenvironment{problem}[1]{\noindent\textbf{#1}:}{\newpage}

\begin{document}

\begin{problem}{1.2}
Transform the expression $3x + 5y - 5z$ into the expression $5(y-z) + 3x$, stating the axiom you used for each step.
\end{problem}


\begin{problem}{1.3}
Let	$\Va = \vecthree{1}{5}{3}$, $\Vb = \vecthree{-5}{2}{6}$, and $\Vc = \vecthree{8}{-7}{-5}$. What is $\Va + \Vb - \Vc$?
\end{problem}

\begin{problem}{1.5}
	Is the function $f(x) = x^2 - x$ linear?
\end{problem}

\begin{problem}{1.6}
	Find the magnitude of the vector $\Vx = \vecthree{6}{-2}{3}$. Use the magnitude to convert $\Vx$ into a unit vector.
\end{problem}

\begin{problem}{1.7}
	Let $\Vx = \vecthree{2}{-3}{1}$ and $\Vy = \vecthree{6}{1}{4}$. What is $\Vx\cdot\Vy$? What is the angle between the vectors?
\end{problem}

\begin{problem}{2.1a}
	Which of the following matrix products are compatible for multiplication?
	\begin{align*}
		\VA &= \Reals^{2\times3} \\
		\VB &= \Reals^{5\times5} \\
		\VC &= \Reals^{3\times5}	
	\end{align*}
Write the dimension of the product whenever the matrices are compatible.

\begin{enumerate}
	\item $\VA\VB$
	\item $\VB\VA$
	\item $\VB\VC$
	\item $\VC\VA$
	\item $\VC\VB$
\end{enumerate}
\end{problem}

\begin{problem}{2.1b}
	Compute the product
	\[ \begin{mex}1&2\\2&1\end{mex} \begin{mex}5&6&3\\1&5&4\end{mex}. \]
\end{problem}

\begin{problem}{2.3}
Simplify the expression $(\VA\VA^{-1})^{-1}\VA$.	
\end{problem}

\begin{problem}{2.6}
	Which is the most efficient way to compute the product $\VA\VB\VC$ when $\dim(\VA) = 100\times400$, $\dim(\VB) = 400\times600$, and $\dim(\VC) = 600\times200$?
	
	\begin{enumerate}
		\item[a.] $(\VA\VB)\VC$
		\item[b.] $\VA(\VB\VC)$
	\end{enumerate}
\end{problem}

\begin{problem}{3.1}
	Rotate the point $\vectwo{3}{-4}$ counter-clockwise by $87\degree$.
\end{problem}

\begin{problem}{3.3}
	Point $\V{p}$ is located at $\vectwo{1}{2}$. Shift $\V{p}$ by 7 horizontally and 6 vertically, and rotate the shifted point $90\degree$ counter-clockwise about the origin. What is the final coordinates of the point?
\end{problem}

\begin{problem}{4.0}
Solve the system of equations
\begin{align*}
	-3x + 5y &= 45 \\
	8x + 6y &= 112	
\end{align*}
\end{problem}


\begin{problem}{5a}
Write a finite difference approximation for the following differential equation using five nodes spanning the interval $x\in[-2,2]$.
\[ e^x\frac{du}{dx} - \sin(x) = 4, \quad u(-2) = 1 \]
Write your equations in matrix form and solve for $u$ at each node.	
\end{problem}


\begin{problem}{5b}
Write a finite difference approximation for the following differential equation using four nodes spanning the interval $t\in[0,1]$.
\[ \frac{d^2y}{dt^2} - t\frac{dy}{dt} = 0, \quad y(0) = 0,\quad y(1) = 4 \]	
\end{problem}

\begin{problem}{6.2}
	Create a $2\times2$ elementary matrix for the row operation $-3R_2\rightarrow R_1$. Use this matrix to perform the row operation on the matrix
	\[ \begin{mex} 2&6\\-1&3\end{mex}. \]	
\end{problem}

\begin{problem}{6.4}
	Use the side-by-side method to compute the inverse of the matrix 
	\[ \begin{mex} 1&2\\3&-2 \end{mex}. \]
\end{problem}

\begin{problem}{7.1a}
	Use Gaussian elimination to compute the rank of the matrix
	\[ \begin{mex} 1&2&-3 \\ -2&-4&6 \\ 0 & 4 & 1 \end{mex}. \]	
\end{problem}

\begin{problem}{7.1b}
	Find a solution to the system of equations
	\begin{align*}
		x_1 + 2x_2 - 3x_3 &= 4 \\
		-2x_1 - 4x_2 + 6x_3 &= -6 \\
		4x_2 + x_3 &= 3	
	\end{align*}
\end{problem}


\end{document}