\documentclass[12pt,ragged]{pajarticle}

%\usepackage{showframe}

\usepackage{lads}

\setlength{\parindent}{0em}
\setlength{\parskip}{1em}

\begin{document}

\AtEndDocument{\section*{Solutions}}

\section*{\hbox{Chapter 11: Vector Spaces, Span, and Basis (Part~I)}}

\begin{question}{1}{%
Are the vectors
\[ \Vv_1 = \vecthree{-1}{0}{2}, \Vv_2 = \vecthree{2}{1}{0}, \Vv_3 = \vecthree{1}{-2}{-1} \]
linearly independent?}{
Let's form a matrix $\VV$ using the vectors as columns.
\[ \VV = \begin{mex} -1 & 2 & 1 \\ 0 & \phan1 & -2 \\ 2 & 0 & -1 \end{mex} \]
This matrix has three rows and three columns. If all three columns are linearly independent, the matrix will have rank three.
\begin{align*}
	\begin{mex} -1 & 2 & 1 \\ 0 & \phan1 & -2 \\ 2 & 0 & -1 \end{mex} \xrightarrow{-R_1}&
	\begin{mex} 1 & -2 & -1 \\ 0 & \phan1 & -2 \\ 2 & 0 & -1 \end{mex} \\
	\xrightarrow{-2R_1\rightarrow R_3}&
	\begin{mex} 1 & -2 & -1 \\ 0 & \phan1 & -2 \\ 0 & 4 & 1 \end{mex} \\
	\xrightarrow{-4R_2\rightarrow R_3}&
	\begin{mex} 1 & -2 & -1 \\ 0 & \phan1 & -2 \\ 0 & 0 & 9 \end{mex} \\
	\xrightarrow{(1/9) R_3}&
	\begin{mex} 1 & -2 & -1 \\ 0 & \phan1 & -2 \\ 0 & 0 & 1 \end{mex}
\end{align*}
The reduced form of \VV\ shows that $\text{rank}(\VV)=3$, so the vectors are linearly independent.
}
\end{question}

\begin{question}{2}{%
Do the vectors $\Vv_1$, $\Vv_2$, and $\Vv_3$ form a basis in $\Reals^3$?
}{
Yes. The space $\Reals^3$ has dimension three, and we have three vectors. We know from the previous question that these three vectors are linearly independent. Combined, these facts prove that the vectors are a basis for $\Reals^3$. Since they are a basis, we also know that the vectors span $\Reals^3$.
}
\end{question}

% -2, 1, 3

\begin{question}{3}{%
Decompose the vector 
\[ \Vu = \vecthree{7}{-5}{-7} \]
onto the vectors $\Vv_1$, $\Vv_2$, and $\Vv_3$.}{
We are looking for scalars $a_1$, $a_2$, and $a_3$ such that
\[ a_1\Vv_1 + a_2\Vv_2 + a_3\Vv_3 = \Vu \]
which is equivalent to solving the linear system
\[ \VV\Va = \Vu \]
where \VV\ is a matrix with columns $\Vv_1$, $\Vv_2$, and $\Vv_3$; and \Va\ is a vector of the scalars $a_1$, $a_2$, and $a_3$:
\[ \begin{mex} -1 & 2 & 1 \\ 0 & \phan1 & -2 \\ 2 & 0 & -1 \end{mex} \vecthree{a_1}{a_2}{a_3} = \vecthree{7}{-5}{-7} \]
We can solve this by Gaussian elimination to find that $a_1=-2$, $a_2=1$, and $a_3=3$.
}	
\end{question}

\begin{question}{4}{%
Are the coefficients you found in Problem 3 the only ones that decompose \Vu\ over the vectors $\Vv_1$, $\Vv_2$, and $\Vv_3$?
}{
Yes. Since the vectors $\Vv_1$, $\Vv_2$, and $\Vv_3$ are a basis, the decomposition of any vector onto them is unique.
}
\end{question}



\end{document}
